Make any stupid plot you want

### 2021-02-15 draft/sections/next.tex
half of whats' next and figures

%"This chapter steps back from the details and considers the larger issues that have been raised. In particular, it considers how this research helps to accomplish the motivating goal of this dissertation, which is the development of computer-based graphical communication." - Mackinlay

\subsection{What have we done}
This work contributes a mathematical description of the transformation from data to visual representation. Combining Butler's fiber bundle model of data with Spivaks formalism of data schemas provides a way of decoupling topology from variability such that the model can support a very large variety of datasets, including discrete relational tables, multivariate high resolution spatio temporal datasets, and complex networks. Modeling the graphic as a fiber bundle provides a way to separate the target display space from the topology of the graphic. By decomposing the mapping from data to visual representation as encoding \vchannel, assembly \vmark, and a mapping between data and graphic topologies \vindex and formalizing what equivariance each stage needs to preserved, this work derives constraints that visualization library authors could embed in their code to guarantee visualizations that are equivariant transforms of the input data.

@@

\subsection{What are the limitations?}

So far this model has only been worked out for a single data set tied to a primitive mark, but it should be extensible to compositing datasets and complex glyphs. The examples and prototype have so far only been implemented for the static 2D case, but nothing in the math limits to 2D and expansion to the animated case should be possible because the model is formalized in terms of the sheaf. While this model supports equivariance of figurative glyphs generated form parameters of the data\cite{beckfeathers2014,byrneFigurativeFramesCritical2017}, it currently does not have a way to evaluate the semantic accuracy of the figurative representation. This model also does not currently factor in effectiveness, but potentially effectiveness criteria could be incorporated into a scheme for assigning encoders and assembling glyphs. This model and the associated prototype is deeply tied to Matplotlib's existing architecture. While the model is expected to generalize to other libraries, such as those built on Mackinlay's APT framework, this has not been worked through. Even though the model is designed to be backend independent, it has only really been tested against the AGG backend. It is especially unknown how this framework interfaces with high performance rendering libraries such as opengl\cite{CarsonOpenGL1997}.

=
\subsection{What else needs to be implemented?}
%%%An interesting control problem will be how to keep the text and graphics consistent. If the rendering exchanges the horizontal and vertical axes of a bar chart, the text must also exchange its references to the axes.
\begin{itemize}
    \item is compose(encodes) and ggplot is plot(encodes, glyph)
    \item heatmaps/ proof of concept of face simplex
    \item composition algebra
    \item boxplot - multiglyph composition
    \item debatable: non-trivial bundle like data on mobius backend
    \item interactivity 
    \item concurrent artist
    \item numpy wrapper
    \item layout
    \item text
    \item keeping all data in sync if transform is changed in one place (so on multi composed graphs, changing x on does it propopgate - part of Bertin invariance)
\end{itemize}
\begin{itemize}
    \item provisional implementation using new architecture in Matplotlib
    \item developer-aimed documentation (API + narrative)
    \item proof of concept 3rd party user-facing library with a biology partner
    \item academic paper about building visualization tools on top of the core architecture
\end{itemize}

>>>>

This work contributes a mathematical description of the transformation from data to visual representation. Combining Butler's fiber bundle model of data with Spivaks formalism of data schemas provides a way of decoupling topology from variability such that the model can support a very large variety of datasets, including discrete relational tables, multivariate high resolution spatio temporal datasets, and complex networks. Modeling the graphic as a fiber bundle provides a way to separate the target display space from the topology of the graphic. By decomposing the mapping from data to visual representation as encoding \vchannel, assembly \vmark, and a mapping between data and graphic topologies \vindex and formalizing what equivariance each stage needs to preserved, this work derives constraints that visualization library authors could embed in their code to guarantee visualizations that are equivariant transforms of the input data.

@@

\subsection{Limitations}

So far this model has only been worked out for a single data set tied to a primitive mark, but it should be extensible to compositing datasets and complex glyphs. The examples and prototype have so far only been implemented for the static 2D case, but nothing in the math limits to 2D and expansion to the animated case should be possible because the model is formalized in terms of the sheaf. While this model supports equivariance of figurative glyphs generated form parameters of the data\cite{beckfeathers2014,byrneFigurativeFramesCritical2017}, it currently does not have a way to evaluate the semantic accuracy of the figurative representation. This model also does not currently factor in effectiveness, but potentially effectiveness criteria could be incorporated into a scheme for assigning encoders and assembling glyphs. This model and the associated prototype is deeply tied to Matplotlib's existing architecture. While the model is expected to generalize to other libraries, such as those built on Mackinlay's APT framework, this has not been worked through. Even though the model is designed to be backend independent, it has only really been tested against the AGG backend. It is especially unknown how this framework interfaces with high performance rendering libraries such as openGL\cite{CarsonOpenGL1997}.

\subsection{Next Steps}
While the model and prototype demonstrate that generation of simple marks from the data, there is a lot of work left to develop a model that underpins a minimally viable library. While the model supports simple composition of glyphs by overlaying glyphs at the same position, more work is needed to define an operator where the fiber bundles have shared $\gbase_2 \hookrightarrow \gbase_1$ such that fibers could be pulled back over the subset. This complex operator is necessary for building semantically meaningful components such as axes labels and interactive visualizations that update on shared \gbase, such as brush-linked views\cite{}

%%%An interesting control problem will be how to keep the text and graphics consistent. If the rendering exchanges the horizontal and vertical axes of a bar chart, the text must also exchange its references to the axes.
\begin{itemize}
    \item is compose(encodes) and ggplot is plot(encodes, glyph)
    \item heatmaps/ proof of concept of face simplex
    \item composition algebra
    \item boxplot - multiglyph composition
    \item debatable: non-trivial bundle like data on mobius backend
    \item interactivity 
    \item concurrent artist
    \item numpy wrapper
    \item layout
    \item text
    \item keeping all data in sync if transform is changed in one place (so on multi composed graphs, changing x on does it propopgate - part of Bertin invariance)
\end{itemize}
\begin{itemize}
    \item provisional implementation using new architecture in Matplotlib
    \item developer-aimed documentation (API + narrative)
    \item proof of concept 3rd party user-facing library with a biology partner
    \item academic paper about building visualization tools on top of the core architecture
\end{itemize}
